Kaggle-in-class Data Challenges can Boost Student Learning

### 2018-03-15 paper-kaggle.Rmd
got title back again, had to change everything

---
title: Kaggle-in-class Data Challenges can Boost Student Learning
# to produce blinded version set to 1
# blinded: 0
author: 
  - name: Julia Polak
    affiliation: Department of Statistics, University of Melbourne
  - name: Dianne Cook
    affiliation: Department of Econometrics and Business Statistics, Monash University
bibliography: bibliography.bib
link-citations: true
fig_caption: yes
keywords: technology, statistical modeling, data science, statistics education
#- that do not appear in the title
abstract: >
  Kaggle is a data modeling competition service, where participants compete to build a model with lower predictive error than other participants. Several years ago they released a reduced service that enables instructors to run competitions in a classroom setting. This paper describes the results of an experiment to determine if the participating in a predictive modeling competition enhances learning. The evidence suggests it does. In addition, students were surveyed to examine if the competition improved engagement and interest in the class.
output: rticles::asa_article
---

@@

# Data collection

This data set was built in two stages. First the original sales in the Melbourne metropole ware downloaded then a few artificial features were add.

The original data was taken from the weekly report published by Domain and compiled by Home Price Guide (R) *COMENT: Found the symbol*. The reports are published as the pdf document. To read them into R we used the XXX package *COMENT: reference to our new package for reading those reports*. The weekly report contains the information about the properties full address, number of bedrooms, property type (house, unit/ duplex, townhouse, development site or other residential), sold price, type of sale (sold, sold prior, passed in, no bid, vendor bid, withdrawn prior to auction, sold after auction and N/A - price or highest bid not available) and agent name. We used the weekly reports from 2nd of February 2013 to 17th of December 2016. We also removed all the records without the price. Over all we had 75,367 sold properties for the data competition.

To enrich this data and make the model building process more interesting we add a few ‘artificial’ features, namely – number of visitors, the average rating gave to the property by the visitors, number of car spaces, number of baths, land size and house size (builded area).

@@

```{r include = FALSE}
# read Unimelb data
UniMelbData <- read.csv(file="data/UniMeldUnIndef.csv", header=TRUE, sep=",")
```

@@

# Add column titles and lay out plots
grid.arrange(grobs=lapply(c(1,5), function(i) {
  arrangeGrob(grobs=pl[i:(i+3)], top=col.titles[i/4 + 1], ncol=1)
}), ncol=2)

Figure \ref{fig:PtqPteBpGroup} shows the boxplots for the $PTQ$ scores and the $PTE$ scores for each group of students[^1].

[^1]: The left boxplot is related to the students that took part in the data competition related to the regression methods, the Melbourne Price competition. The right boxplot is related to the students that took part in the data competition related to the calssification methods, the Spam classification competition.

@@

To examine the correlation between student’s engagement levels and the performances in the exam we plot the number of submission during the competition versus the performances in the exam, Figure \ref{fig:numsubmition}. Once again we exam the performances based on two normalizations. Once normalizing by the total possible marks for the relevant cluster of questions[^2] ($PTQ$) and once normalizing by total exam marks ($PTE$).

[^2]: For the students participated in the Melbourne Price competition is the cluster of regression questions. For the students participated in the Spam Classification competition is the cluster of questions about classification methods.

@@

>>>>

---
title: Kaggle-in-class Data Challenges can Boost Student Learning

# to produce blinded version set to 1
blinded: 0

authors: 
  - name: Julia Polak
    affiliation: Department of Statistics, University of Melbourne
  - name: Dianne Cook
    affiliation: Department of Econometrics and Business Statistics, Monash University

keywords:
- instructional technology
- statistical modeling
- data science
- statistics education
- data mining

abstract: |
  Kaggle is a data modeling competition service, where participants compete to build a model with lower predictive error than other participants. Several years ago they released a reduced service that enables instructors to run competitions in a classroom setting. This paper describes the results of an experiment to determine if the participating in a predictive modeling competition enhances learning. The evidence suggests it does. In addition, students were surveyed to examine if the competition improved engagement and interest in the class.

bibliography: bibliography.bib
output: rticles::asa_article
---

@@

@@

```{r}
# read Unimelb data
UniMelbData <- read.csv(file="data/UniMeldUnIndef.csv", header=TRUE, sep=",")
```

@@

# Add column titles and lay out plots
grid.arrange(grobs=lapply(c(1,5), function(i) {
  arrangeGrob(grobs=pl[i:(i+3)], top=col.titles[i/4 + 1], ncol=1)
}), ncol=2)
```

Figure \ref{fig:PtqPteBpGroup} shows the boxplots for the $PTQ$ scores and the $PTE$ scores for each group of students. The left boxplot is related to the students that took part in the data competition related to the regression methods, the Melbourne Price competition. The right boxplot is related to the students that took part in the data competition related to the calssification methods, the Spam classification competition.

@@

To examine the correlation between student’s engagement levels and the performances in the exam we plot the number of submission during the competition versus the performances in the exam, Figure \ref{fig:numsubmition}. Once again we exam the performances based on two normalizations. Once normalizing by the total possible marks for the relevant cluster of questions ($PTQ$) and once normalizing by total exam marks ($PTE$). For the students participated in the Melbourne Price competition is the cluster of regression questions. For the students participated in the Spam Classification competition is the cluster of questions about classification methods.

@@

# Stuff to hang onto for now

\section{Verifications}
\label{sec:verify}
This section will be just long enough to illustrate what a full page of
text looks like, for margins and spacing.

# Appendix

There were two data sets used for competitions, one was a regression problem and the other a classification problem.

This data contained records of auction prices for residentia properties in Melbourne. The data was compiled from auction reports collected between Feb 2, 2013 and Dec 17, 2016.

Auction reports were published weekly by Domain and compiled by Home Price Guide (R) *COMMENT: Found the symbol* in pdf format. To read them into R we used the XXX package *COMMENT: reference to our new package for reading those reports*. The reports contained the following information about the properties sold during the week: full address, number of bedrooms, property type (house, unit/ duplex, townhouse, development site or other residential), sold price, type of sale (sold, sold prior, passed in, no bid, vendor bid, withdrawn prior to auction, sold after auction and N/A - price or highest bid not available) and agent name. We also removed all the records without the price. Over all we had 75,367 sold properties for the data competition.

To enrich this data, to make it more realistic we added several additional features, number of visitors, the average rating given to the property by the visitors, number of car spaces, number of baths, land size and house size.	It is possible that some of this information might have been extracted from other web sites, given that we had the full address of the property, but the time required to research that was too daunting to tackle for the upcoming teaching period.
